Effect of quality-adjusted production function inputs

### 2015-05-17 Articles/Presentation_ESEE_2015/ESEE_2015.Rnw
Working on dynamic solow residual.

<<Data>>=
# Create the historicalData data frames
historicalData <- list()
historicalData$PT <- filter(IST, Country == "PT")
historicalData$UK <- filter(Leeds, Country == "UK")
allHistoricalData <- do.call(plyr::rbind.fill, historicalData)
@

@@

Unadjusted and quality-adjusted data for both Portugal and the United Kingdom 
are shown in Figure~\ref{fig:Historical Data}.

<<Historical Data, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Historical data.", warning=FALSE>>=
histDataForGraph <- leaf_apply(formulas, class = "formula", f=function(form, name){
  meta <- strsplit(name, split="[.]") %>% unlist
  Country <- meta[1]
  flavor <- meta[2]
  measure.vars <- c(summands(form, right=FALSE), summands(form, 1:3))
  trimmed_historical_data <- historicalData[[Country]][c("Year", measure.vars)]
  mtrimmed_historical_data <- melt(trimmed_historical_data, measure.vars = measure.vars, 
                                   value.name = "value", variable.name = "variable")
  mtrimmed_historical_data$Country <- Country
  mtrimmed_historical_data$flavor <- if (flavor == "adjusted") "Quality-adjusted" else "Unadjusted"
  return(mtrimmed_historical_data)
}) %>% do.call(rbind, .)

@@

>>>>

<<Data>>=
# Create the historicalData data frames
historicalData <- list()
commonVars <- c("Year", "Country", "Source")
historicalData$PT <- filter(IST, Country == "PT") 
historicalData$PT <- historicalData$PT[ , unique(c(commonVars,
                                                   all.vars(formulas$PT$unadjusted), 
                                                   all.vars(formulas$PT$adjusted)))]
historicalData$UK <- filter(Leeds, Country == "UK")
historicalData$UK <- historicalData$UK[ , unique(c(commonVars,
                                                   all.vars(formulas$UK$unadjusted), 
                                                   all.vars(formulas$UK$adjusted)))]
# allHistoricalData <- do.call(plyr::rbind.fill, historicalData)
@

@@

Unadjusted and quality-adjusted data for both Portugal and the United Kingdom 
are shown in Figure~\ref{fig:Historical Data Graph}.

<<Historical Data Graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Historical data.", warning=FALSE>>=
# Melt the data into a form that is useful in a couple places
histDataForGraph <- leaf_apply(formulas, class = "formula", f=function(form, name){
  meta <- strsplit(name, split="[.]") %>% unlist
  Country <- meta[1]
  flavor <- meta[2]
  measure.vars <- c(summands(form, right=FALSE), summands(form, 1:3))
  trimmed_historical_data <- historicalData[[Country]][c("Year", measure.vars)]
  mtrimmed_historical_data <- melt(trimmed_historical_data, measure.vars = measure.vars, 
                                   value.name = "value", variable.name = "variable")
  mtrimmed_historical_data$Country <- Country
  mtrimmed_historical_data$flavor <- if (flavor == "adjusted") "Quality-adjusted" else "Unadjusted"
  return(mtrimmed_historical_data)
}) %>% do.call(rbind, .)

@@

%++++++++++++++++++++++++++++++
\subsection{Dynamic Solow Residual} 
\label{sec:dynamic_solow_residual}
%++++++++++++++++++++++++++++++

This section presents results of dynamic Solow Residual analyses.

<<Dynamic Solow Residual>>=
# Calculate finite difference approximations of time derivatives for output and all factors of production
lag <- 1       # Desired lag for calculations. Set to n for differences across n years.
pad <- "tail"  # Desired padding for NA values

fd <- leaf_apply(formulas, class = "formula", f = function(form, name){
  meta <- strsplit(name, split="[.]") %>% unlist
  Country <- meta[1]
  flavor <- meta[2]

  denom <- summands(form, n=length(summands(form))) # time variable
  nums <- c(summands(form, n=1, right = FALSE), summands(form, n=1:3)) # y, k, l, and e
  trimmedData <- historicalData[[Country]][c(denom, nums)] # columns are: iYear, y, k, l, and e
  df <- list()
  df <- historicalData[[Country]][ , c("Year", "Country")]
  df$dydt <- ediff(trimmedData[[2]], pad=pad, lag=lag) / ediff(trimmedData[[1]], pad=pad, lag=lag)
  df$dkdt <- ediff(trimmedData[[3]], pad=pad, lag=lag) / ediff(trimmedData[[1]], pad=pad, lag=lag)
  df$dldt <- ediff(trimmedData[[4]], pad=pad, lag=lag) / ediff(trimmedData[[1]], pad=pad, lag=lag)
  df$dedt <- ediff(trimmedData[[5]], pad=pad, lag=lag) / ediff(trimmedData[[1]], pad=pad, lag=lag)

  df$flavor <- flavor
  return(df)
}) %>% do.call(rbind, .)

# origCDKLCoeffs <- filter(AllCoeffs, method=="orig", model=="cd")
# origCDKLCoeffs <- origCDKLCoeffs %>% select(method, country, Source, nestStr, energy, lambda, alpha, beta, gamma)
# # Change name of Source REXS to REXS1960 and relevel to conform to information in fdlambda.
# origCDKLCoeffs$Source <- as.character(origCDKLCoeffs$Source)
# origCDKLCoeffs$Source <- replace(origCDKLCoeffs$Source, which(origCDKLCoeffs$Source=="REXS"), "REXS1960")
# origCDKLCoeffs$Source <- as.factor(origCDKLCoeffs$Source)
# origCDKLCoeffs$Source <- relevelFactor(origCDKLCoeffs$Source, levs = c("REXS1960", "IST", "Leeds"))
# 
# # Fill new columns in the fd data frame with 
# # lambda, alpha, beta, and gamma values from the parameter estimation process.
# # lambda, alpha, beta, and gamma values for the orig fits are in the origCDKLCoeffs data frame.
# # Using the hint at 
# # http://stackoverflow.com/questions/15059076/
# # r-how-to-call-apply-like-function-on-each-row-of-dataframe-with-multiple-argum
# labg <- apply(fd[ , c('Country', 'Source', 'energyType')], 1, function(x){
#     row <- as.data.frame(filter(origCDKLCoeffs, 
#                                 country==x["Country"], Source==x["Source"], energy==x["energyType"]))
#     abg <- subset(row, select=c(lambda, alpha, beta, gamma))
#     if (nrow(abg) < 1){ abg[1, ] <- NA }
#     return(abg)
#   })
# cols <- do.call(rbind, labg)
# fd <- cbind(fd, cols)
# 
# # Now calculate lambda by finite difference for every year to make a time series
# fdlambda <- mutate(fd, lambda_diff = dydt/iGDP - (alpha/iK*dkdt + beta/iL*dldt + gamma/iE*dedt))
# 
# # Calculate the fraction of GDP growth in each year that is explained by the solow residual
# fdlambda <- mutate(fdlambda, solowFrac = lambda_diff / (dydt/iGDP))
# 
# # Calculate average values for lambda and solowFrac
# avgs <- fdlambda %>% 
#   group_by(Source, Country, energyType) %>% 
#   summarise(
#     avg.lambda_diff = mean(lambda_diff, na.rm=TRUE),
#     avg.solowFrac = mean(solowFrac, na.rm=TRUE)
#     )

@
